\documentclass[letterpaper]{article}
% Required Packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
%%%%%%%%%%
% PDFINFO for PDFL A TEX
% Uncomment and complete the following for metadata
%(your paper must compile with PDFL A TEX)
\pdfinfo{
/Title (Input Your Paper Title Here)
/Author (John Doe, Jane Doe)
/Keywords (Input your paper’s keywords in this optional area)
}%%%%%%%%%%
% Section Numbers
% Uncomment if you want to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
%%%%%%%%%%
% Title, Author, and Address Information
\title{A Scrabble Agent and Whatnot}
\author{Casey Pore \\
Colorado State University\\
cpore@rams.colostate.edu\\
}
%%%%%%%%%%
% Body of Paper Begins
\begin{document}
\maketitle

\section{Problem Description}
what were you trying to do; this should describe the problem addressed by the project. Use your original proposals as a starting point for this part. Also, include your hypothesis(es) here along with a precis of what you found.

\section{Previous Work}
survey what approaches have been adopted before. It may be a quick reference to the AI techniques selected for the project or it could reference papers on the problem or on the previous approaches taken to the problem. It depends on the composition of the project; if it extends previous work, then reference what it extends; if it is novel, then reference techniques that motivated the design. References should be from publications rather than web sites; the exception to that is if you downloaded code or problems from a website.

\section{Approach Taken}
what did you do. Describe your project: what AI techniques are used by it, why you picked these techniques, how was the project structured, who did what, what sort of data was supplied (example problems for learning systems, prior models for non-learners), what results were expected. Code samples must be short.

\section{Evaluation and Analysis}
How well did your project do: as expected, better or worse. Why did it perform as it did? What worked and what did not? Were there any surprises? What experiments/evaluation did you run? Include your experiment design to this section; it may have a subsection for the experiment design and another for the analysis.

\section{Future Work, Conclusions}
What did you learn? What would you do in addition or differently? \cite{Torre2008}



% References and End of Paper
\bibliography{Scrabble}
\bibliographystyle{aaai}
\end{document}